### 2012-05-23 doc/hmcsa_adjoint_vs_direct/direct_vs_adjoint.tex
Updated tech note for direct vs. adjoint study

Fig. \ref{fig:cpu_time} gives the CPU time needed for each case to
converge in seconds and Fig. \ref{fig:iterations} gives the number of
iterations needed for each case to converge.

%%---------------------------------------------------------------------------%%
\section{Convergence Study Results}

%%---------------------------------------------------------------------------%%
\section{Conclusion}

>>>>

Fig. \ref{fig:cpu_time} gives the CPU time needed for each case to
converge in seconds and Fig. \ref{fig:iterations} gives the number of
iterations needed for each case to converge. We see clearly in
Fig. \ref{fig:cpu_time} that the using the adjoint solver with MCSA
results in several orders of magnitude speedup over the direct solver
while the number of iterations required to converge is of a similar
scale. We expect this for several reasons. First, with an equivalent
number of histories specified for both solvers and a system of size
$N$, the direct solver will compute $N$ times the number of histories
random walks while the adjoint solver will only compute the number of
histories specified. Initially, this may not seem like a fair
comparison, however, we see from Fig. \ref{fig:cpu_time} that the
number of iterations requried to converge is approximately the
same. This is necessary in the direct method to ensure a contribution
from each state as the random walk sequence will only contribute to
the starting state. For the adjoint method, because the random
sequence contributes to the state in which it currently resides, fewer
histories are neccessary to compute contributions from all the nonzero
parts of the source.

%%---------------------------------------------------------------------------%%
\section{Convergence Study Results}
Although CPU time was shown to be greatly decreased by using the
adjoint solver in the preceeding section, the iteration behavior was
approximately the same. To assess the convergence properties of MCSA
using each solver and stencil, the infinity norm of the residual
computed in Eq. \ref{eq:mcsa_step2} was computed at each iteration for
a fixed problem size. Fig. \ref{fig:convergence} gives the results of
these computuations. First, it is worthy to note on the semilog plot
that we are indeed achieving the expected exponential convergence from
MCSA. Second, we note that both the direct and adjoint methods have
approximately equivalent convergence behavior. Finally, it is observed
the the fourth-order accurate Laplacian stencil converges in fewer
iterations than the second-order accurate stencil as expected.

%%---------------------------------------------------------------------------%%
\section{Conclusion}
A parameter study has been completed to verify that an adjoint
stochastic solver should be used with an MCSA implementation instead
of a direct stochastic solver. Several orders of magnitude reduction
in CPU time is observed when using the adjoint solver over the
direct. In addition, it was observed that MCSA had similar exponential
convergence properties with both the direct and adjoint solvers. Even
for cases where fewer iterations were needed to converge to a solution
using the direct solver, the large number of histories required caused
the CPU time to become large compared to the adjoint method.
